Refunctionalization of Abstract Abstract Machines (Functional Pearl)

### 2018-06-02 paper/paper.tex
update
update the code in paper

\begin{lstlisting}
case Let(x, App(f, ae), e) =>
  val closures = atomicEval(f, env, store).asInstanceOf[Set[Clos]]
  nd[Clos, Ans](closures, Set[VS](), { case (clos, accouter, closnd) =>
    val Clos(Lam(v, body), c_env) = clos
    val baddr = allocBind(v, new_time)
    val new_env = c_env + (v -> baddr)
    val new_store = store.update(baddr, atomicEval(ae, env, store))
    aeval(body, new_env, new_store, new_time, (bodyvss: Set[VS]) => {
      nd[VS, Ans](bodyvss, Set[VS](), { case (vs, accinner, bdnd) =>
        val VS(vals, time, store) = vs
        val new_env = env + (x -> baddr)
        val new_store = store.update(baddr, vals)
        aeval(e, new_env, new_store, time,
              /* accumulate the values of one path, and call next body value*/
              (evss: Ans) => bdnd(accinner ++ evss))
      },
      /* accumulate the values of paths to accouter, and call next closure */
      (evss: Ans) => closnd(evss ++ accouter))
    })
  },
  continue)
\end{lstlisting}

The outer call to @nd@ evaluates over the set of possible closures of
@f@. For each such closure, we go into its body with a new environment and a new store.
Remember that the function @aeval@ returns a set of @VS@ objects to its continuation,
so the continuation argument @bodyvss@ is a set of @VS@ objects which represent values
from multiple computation paths when evaluating this single @body@ expression.

We next consider the inner call to @nd@, which evaluates over the set of body values @bodyvss@.
For each @VS@, the timestamp and store in @VS@ will be instantiated to
@aeval@.
The inner application of @aeval@ returns a set of values of one computation path to its
continuation, where the continuation will accumulate the result and transfer
evaluation to the next body value.

@@

\begin{lstlisting}
def aeval(e: Expr, env: Env, store: BStore, time: Time, cache: Cache, continue: Cont): Ans = {
  val config = Config(e, env, store, time)
  /* lookups the out cache, uses the values if contains this config */
  if (cache.outContains(config)) return continue(Ans(cache.outGet(config), cache))
  /* uses the values from in cache */
  val new_cache = cache.outUpdate(config, cache.inGet(config))
  val new_time = (e::time).take(k)
  e match {
    case Let(x, App(f, ae), e) =>
      val closures = atomicEval(f, env, store).asInstanceOf[Set[Clos]]
      nd[Clos, Ans](closures, Ans(Set[VS](), new_cache), { case (clos, Ans(accouter, cache), closnd) =>
        ......
        aeval(body, new_env, new_store, new_time, cache, { case Ans(bodyvss, bodycache) =>
          nd[VS, Ans](bodyvss, Ans(Set[VS](), bodycache), { case (vs, Ans(acc_vss, cache), bdnd) =>
            ......
            aeval(e, new_env, new_store, time, cache, { case Ans(evss, ecache) =>
              bdnd(Ans(accinner ++ evss, ecache))
            })
          },
          { case Ans(evss, cache) => closnd(Ans(evss ++ accouter, cache)) })
        })
      },
      /* updates the out cache after evaluation */
      { case Ans(vss, cache) => continue(Ans(vss, cache.outUpdate(config, vss))) })

    case ae if isAtomic(ae) =>
      val vs = Set(VS(atomicEval(ae, env, store), new_time, store))
      continue(Ans(vs, new_cache.outUpdate(config, vs))) /* updates the out cache after evaluation */
  }
}
\end{lstlisting}

After completing the evaluation, we obtain a set of @VS@ objects and must update the @out@ cache.
In the first case of pattern matching, this happens inside the last continuation of the
outer @nd@ call.
For the second case, we update the @out@ cache before calling the @continue@ continuation.

@@

\begin{lstlisting}
def aeval(e: Expr, env: Env, store: BStore, time: Time, cache: Cache): Ans @cps[Ans] = {
  val config = Config(e, env, store, time)
  if (cache.outContains(config)) Ans(cache.outGet(config), cache)
  else {
    val new_time = (e::time).take(k)
    val new_cache = cache.outUpdate(config, cache.inGet(config))
    e match {
      case Let(x, App(f, ae), e) =>
        val closures = atomicEval(f, env, store).asInstanceOf[Set[Clos]]
        val (Clos(Lam(v, body), c_env), clscache) = ndcps[Clos](closures, Ans(Set[VS](), new_cache))
        val vbaddr = allocBind(v, new_time)
        val new_cenv = c_env + (v -> vbaddr)
        val new_cstore = store.update(vbaddr, aeval(ae, env, store))
        /* evaluates the function application App(f, ae) */
        val Ans(bodyvss, bodycache) = aeval(body, new_cenv, new_cstore, new_time, clscache)
        val (VS(vals, time, vsstore), vscache) = ndcps[VS](bodyvss, Ans(Set[VS](), bodycache))
        val baddr = allocBind(x, time)
        val new_env = env + (x -> baddr)
        val new_store = vsstore.update(baddr, vals)
        /* evaluates the body expression from the let */
        val Ans(finval, fincache) = aeval(e, new_env, new_store, time, vscache)
        Ans(finval, fincache.outUpdate(config, finval))

      case ae if isAtomic(ae) =>
        val vs = Set(VS(atomicEval(ae, env, store), new_time, store))
        Ans(vs, cache.outUpdate(config, vs))
    }
  }
}
\end{lstlisting}

@@

%\section{Perspective}
%\subsection{Defunctionalization}
%specify higher-order definitional interpreters using first-order means (cite)
%\subsection{Refunctionalization}
%\section{Detour: Shiver's \textit{k}-CFA?}

>>>>

\begin{lstlisting}
case Let(x, App(f, ae), e) =>
  val closures = aeval(f, env, store).asInstanceOf[Set[Clos]]
  nd[Clos, Ans](closures, Set[VS](), { case (clos, clsacc, clsnd) =>
    val Clos(Lam(v, body), c_env) = clos
    val baddr = allocBind(v, new_time)
    val new_cenv = c_env + (v -> baddr)
    val new_store = store.update(baddr, aeval(ae, env, store))
    aeval(body, new_cenv, new_store, new_time, (bdvss: Set[VS]) => {
      nd[VS, Ans](bdvss, Set[VS](), { case (vs, bdacc, bdnd) =>
        val VS(vals, time, store) = vs
        val baddr = allocBind(x, time)
        val new_env = env + (x -> baddr)
        val new_store = store.update(baddr, vals)
        aeval(e, new_env, new_store, time, { 
          /* accumulate the values of one path, and call next body value */
          case evss => bdnd(bdacc ++ evss) 
        })
      }, { 
        /* accumulate the values of paths, and call next closure */
        case evss => clsnd(evss ++ clsacc) 
      })
    })
  },
  continue)
\end{lstlisting}

The outer call to @nd@ evaluates over the set of possible closures of
@f@. For each such closure, we go into its body with a new environment and a new store.
Remember that the function @aeval@ returns a set of @VS@ objects to its continuation,
so the continuation argument @bdvss@ is a set of @VS@ objects which represent values
from multiple computation paths when evaluating this single @body@ expression.

We next consider the inner call to @nd@, which evaluates over the set of body values @bdvss@.
For each @VS@, the timestamp and store in @VS@ will be instantiated to
@aeval@.
The inner application of @aeval@ returns a set of values of one computation path to its
continuation, where the continuation will accumulate the result and transfer
evaluation to the next body value.

@@

\begin{lstlisting}
def aeval(e: Expr, env: Env, store: BStore, time: Time, cache: Cache, continue: Cont): Ans = {
  val config = Config(e, env, store, time)
  /* lookups the out cache, uses the values if contains this config */
  if (cache.outContains(config)) cont(Ans(cache.outGet(config), cache))
  else {
    val new_cache = cache.outUpdate(config, cache.inGet(config))
    val new_time = (e::time).take(k)
    e match {
      case Let(x, App(f, ae), e) =>
        val closures = atomicEval(f, env, store).asInstanceOf[Set[Clos]]
        nd[Clos, Ans](closures, Ans(Set[VS](), new_cache), { 
          case (Clos(Lam(v, body), c_env), clsans, clsnd) =>
            ...... /* sets vairables, e.g., new_cenv and new_store */
            aeval(body, new_cenv, new_store, new_time, clsans.cache, { case Ans(bdvss, bdcache) =>
                nd[VS, Ans](bdvss, Ans(Set[VS](), bdcache), { 
                  case (VS(vals, time, vssstore), bdans, bdnd) =>
                    ...... /* sets variables, e.g., new_env and new_store */
                    aeval(e, new_env, new_store, time, bdans.cache, { 
                      /* updates the out cache after evaluated to the end of one path  */
                      case Ans(evss, ecache) => bdnd(bdans ++ Ans(evss, ecache.outUpdate(config, evss)))
                    })
                }, { case eans => clsnd(eans ++ clsans) })
            })
        }, 
        cont)

      case ae if isAtomic(ae) =>
        val vs = Set(VS(atomicEval(ae, env, store), new_time, store))
        cont(Ans(vs, new_cache.outUpdate(config, vs)))
    }
  }
}
\end{lstlisting}

After completing the evaluation, we obtain a set of @VS@ objects and must update the @out@ cache.
In the first case of pattern matching, this happens at the innermost continuation of @aeval@,
since it is where we reach the end of one computation path.
For the second case, we update the @out@ cache before calling the @continue@ continuation.

@@

\begin{lstlisting}
def aeval(e: Expr, env: Env, store: BStore, time: Time, cache: Cache): Ans @cps[Ans] = {
  val config = Config(e, env, store, time)
  if (cache.outContains(config)) Ans(cache.outGet(config), cache)
  else {
    val new_time = (e::time).take(k)
    val new_cache = cache.outUpdate(config, cache.inGet(config))
    e match {
      case Let(x, App(f, ae), e) =>
        val closures = atomicEval(f, env, store).asInstanceOf[Set[Clos]]
        val (Clos(Lam(v, body), c_env), clscache) = ndcps[Clos](closures, Ans(Set[VS](), new_cache))
        val vbaddr = allocBind(v, new_time)
        val new_cenv = c_env + (v -> vbaddr)
        val new_store = store.update(vbaddr, atomicEval(ae, env, store))
        /* evaluates the function application App(f, ae) */
        val Ans(bdvss, bdcache) = aeval(body, new_cenv, new_store, new_time, clscache)
        val (VS(vals, time, vsstore), vscache) = ndcps[VS](bdvss, Ans(Set[VS](), bdcache))
        val baddr = allocBind(x, time)
        val new_env = env + (x -> baddr)
        val new_vsstore = vsstore.update(baddr, vals)
        /* evaluates the body expression from the let */
        val Ans(evss, ecache) = aeval(e, new_env, new_vsstore, time, vscache)
        Ans(evss, ecache.outUpdate(config, evss))

      case ae if isAtomic(ae) =>
        val vs = Set(VS(atomicEval(ae, env, store), new_time, store))
        Ans(vs, new_cache.outUpdate(config, vs))
    }
  }
}
\end{lstlisting}

@@

%specify higher-order definitional interpreters using first-order means (cite)
%\section{Detour: Shiver's \textit{k}-CFA?}
